Fisher vector places: learning compact descriptors for place recognition

### 2014-11-14 cvpr15review.tex
results section

%%%%%%%%% ABSTRACT
\begin{abstract}
   The goal of this work is to localize a query photograph by finding other images depicting the same place in a large geotagged image database. The contributions of this work are three fold. First, we learn a {\em discriminative} yet {\em compact} descriptor of each image in the database. This is achieved by applying exemplar support vector machine (e-SVM) learning to compact Fisher vector descriptors extracted from database images. Secondly, we analyze the exemplar support vector machine objective and show that the learnt hyperplane can be interpreted as a new descriptor that replaces the original positive example and is re-weighted to increase its separation from the negative data. Finally, based on this analysis we demonstrate that the learnt and re-normalized descriptor could be directly used for matching, thus avoiding the need for expensive and tedious calibration typically needed for exemplar support vector machine methods. Place recognition results are shown on two image datasets of Google street-view images from Pittsburgh  and demonstrate the learnt representation consistently improves over the standard Fisher vector descriptors at different target dimensions. 
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}

The goal of this work is to localize a query image by matching to a large database of geotagged street-level imagery.
This is an important problem with practical applications in robotics, augmented reality or navigation. This task is however very difficult. It is hard to distinguish different places, e.g.\ streets in a city, from each other. The imaged appearance of a place can change drastically due to factors such as  viewpoint, illumination or even changes over time.
%
Finally, with the emergence of planet-scale geotagged image collections, such as Google Street-view, the image databases are becoming very large. We estimate a single country like France is covered by more than 60 million street-level panoramas. Hence the fundamental challenge in place recognition lies now in designing robust, discriminative, yet compact, image representations.

@@

         In detail, for $w=w^{\perp}+w^{||}$ the objective~(\ref{eq:obj}) can be written as
            %
            \begin{align}
               \label{eq:decomposedOrig}
                 ||w^{\perp}+w^{||}||^{2} &+
                 C_1 \cdot h
                 \left(
                  (w^{\perp}+w^{||})^T\Phi^+ + b
                 \right) \\ \nonumber
                 &+
                 C_2\sum_{\Phi\in \mathcal N} h
                 \left(
                  -(w^{\perp}+w^{||})^T\Phi-b
                 \right).
            \end{align}
            %
         Note that the orthogonal part $w^{\perp}$ does not change the value of  the second term in~(\ref{eq:decomposedOrig}) because $(w^{\perp}+w^{||})^T\Phi^+=(w^{||})^T\Phi^+$, and hence~(\ref{eq:decomposedOrig}) reduces to
            %
            \begin{align} 
               \label{eq:decomposed} 
               ||w^{\perp}+w^{||}||^{2} &+
               C_1 \cdot h
               \left(
                   w^{||T}\Phi^+ +b_j
               \right) \\ \nonumber
               &+
               C_2\sum_{\Phi\in \mathcal N} h
               \left(
                   -(w^{\perp}+w^{||})^T\Phi-b
               \right).
            \end{align}
            %
         In the limit case as $C_2 \rightarrow 0$ any non-zero component $w^{\perp}$
         will increase the value of the objective~\eqref{eq:decomposed}. This can be seen by noting that the third term vanishes when $C_2 \rightarrow 0$ and hence the objective is dominated by the first two terms. Further, the second term in~\eqref{eq:decomposed} is independent of $w^{\perp}$. Finally, the first term will always increase for any non-zero value of $w^{\perp}$ as $||w^{\perp}+w^{||}||^{2} \geq ||w^{||}||$ for any $w^{\perp}\neq0$.

@@

      \paragraph{Bag-of-visual-words}
         \textcolor{petr}{
            For bag of visual words image representation we provide three kind of results summarized in table \ref{tab:recallBOW}. First, similarly to \cite{Gronat13}, we compare proposed method (e-SVM) to the standard bag-of-visual-words baseline (BOW). We observe improvement from 28.7\% recall@K=1 (BOW) to 31.8\% recall@K=1 (BOW e-SVM). Secondly, the results are also confronted to the p-value calibration of \cite{Gronat13}. The p-value calibration (BOW p-val) gives marginally better results (33.6\% recall@K=1) but at an increased training cost and a significantly higher required memory footprint at query time, which makes the method hard to scale-up to database sizes beyond 200k images. Finally, we investigate the performance of a raw exemplar-SVM (raw e-SVM) with no calibration. As reported in \cite{Gronat13}, we observe that using these scores directly with no calibration does not perform on BOW. However, this method gives surprisingly good results on Fisher vectors. Details will be given next.
         }

         \begin{table}[b]
            \begin{centering}
               \include{table25kBOW}
            \caption{ \textbf{Results on Pittsburgh 25k dataset - BOW.}
                  The fraction of correctly recognized queries (recall@K)
                  \textcolor{petr}{
                     in the top $K\in\{1,2,5,10,20\}$ retrieved images.
                     % Different methods have been applied to two types of descriptors, the BOW and Fisher vectors compressed to different dimensions. The learnt representations (BOW \emph{e-SVM} and BOW \emph{p-val}) outperform the raw BOW baseline as well as the learnt representation without calibration (BOW raw SVM).
                  } 
                  The learnt descriptors by the proposed method (BOW e-SVM) marginally improves over the raw bag-of-visual-words descriptors (BOW) across the whole range of $K$  and all dimensions.
%                   on both the 25k and 55k Pittsburgh image datasets.     
            }
            \label{tab:recallBOW}
            \end{centering}
         \end{table}

      \paragraph{Fisher vectors}

@@

         Figure \ref{fig:recall} shows the results in a form of curve and it clearly demonstrates the benefits of the learnt descriptors. For comprehensiveness the figure also shows the BOW baseline curve from table \ref{tab:recallBOW} shown in black.

@@

         The p-val calibration of~\cite{Gronat13} has been tested on e-SVM Fisher vectors. We found that the calibration did not perform well for the learnt Fisher vector classifiers (top 1 recall of 25.3\% compared to baseline performance of 33.6\% for dimension 128). When examining the results, we believe this may be attributed to the fact that Fisher vectors are low dimensional compared to the extremely high-dimensional (d=100,000) bag-of-visual-words representation, which affects the distribution of the classifier scores on the negative test data.

@@

         Next, we investigate the performance of a raw exemplar-SVM (raw e-SVM) without any further calibration for both image descriptors . We observe that while using raw SVM scores directly without any calibration fails for the bag-of-visual-words representation (BOW raw SVM) (as reported in \cite{Gronat13}), it marginally improves recall for larger $K$ for Fisher vectors (68.4\% for FV128 raw SVM compared to 67.7\% for non-learned FV128).

@@

         Finally, we performed experiments on the Pittsburgh 55k dataset. Similarly to the former dataset, the results summarized in table \ref{tab:recall2} show consistent improvement of the proposed method (FV e-SVM) over the Fisher vector baseline (FV) across all target dimensions. The benefits of the proposed method in terms of memory efficiency will be discussed next.

>>>>

%%%%%%%%% ABSTRACT
\begin{abstract}
   \noindent
   The goal of this work is to localize a query photograph by finding other images depicting the same place in a large geotagged image database. The contributions of this work are three fold. First, we learn a {\em discriminative} yet {\em compact} descriptor of each image in the database. This is achieved by applying exemplar support vector machine (e-SVM) learning to compact Fisher vector descriptors extracted from database images. Secondly, we analyze the exemplar support vector machine objective and show that the learnt hyperplane can be interpreted as a new descriptor that replaces the original positive example and is re-weighted to increase its separation from the negative data. Finally, based on this analysis, we demonstrate that the learnt and re-normalized descriptor could be directly used for matching, thus avoiding the need for expensive and tedious calibration typically needed for exemplar support vector machine methods. Place recognition results are shown on two image datasets of Google street-view images from Pittsburgh  and demonstrate the learnt representation consistently improves over the standard Fisher vector descriptors at different target dimensions. 
\end{abstract}

%%%%%%%%% BODY TEXT
\section{Introduction}
\noindent
The goal of this work is to localize a query image by matching to a large database of geotagged street-level imagery.
This is an important problem with practical applications in robotics, augmented reality or navigation. This task is however very difficult. It is hard to distinguish different places, e.g.\ streets in a city, from each other. The imaged appearance of a place can change drastically due to factors such as  viewpoint, illumination or even changes over time.
%
Finally, with the emergence of planet-scale geotagged image collections, such as Google Street-view, the image databases are becoming very large. We estimate a single country like France is covered by more than 60 million street-level panoramas. Hence the fundamental challenge in place recognition lies now in designing robust, discriminative, yet compact, image representations.

@@

         In detail, for $w=w^{\perp}+w^{||}$, the objective~(\ref{eq:obj}) can be written as
            %
            \begin{align}
               \label{eq:decomposedOrig}
                 ||w^{\perp}+w^{||}||^{2} &+
                 C_1 \cdot h
                 \left(
                  (w^{\perp}+w^{||})^T\Phi^+ + b
                 \right) \\ \nonumber
                 &+
                 C_2\sum_{\Phi\in \mathcal N} h
                 \left(
                  -(w^{\perp}+w^{||})^T\Phi-b
                 \right).
            \end{align}
            %
         Note that the orthogonal part $w^{\perp}$ does not change the value of  the second term in~(\ref{eq:decomposedOrig}) because $(w^{\perp}+w^{||})^T\Phi^+=(w^{||})^T\Phi^+$, and hence~(\ref{eq:decomposedOrig}) reduces to
            %
            \begin{align} 
               \label{eq:decomposed} 
               ||w^{\perp}+w^{||}||^{2} &+
               C_1 \cdot h
               \left(
                   w^{||T}\Phi^+ +b_j
               \right) \\ \nonumber
               &+
               C_2\sum_{\Phi\in \mathcal N} h
               \left(
                   -(w^{\perp}+w^{||})^T\Phi-b
               \right).
            \end{align}
            %
         In the limit case as $C_2 \rightarrow 0$ any non-zero component $w^{\perp}$
         will increase the value of the objective~\eqref{eq:decomposed}. This can be seen by noting that the third term vanishes when $C_2 \rightarrow 0$ and hence the objective is dominated by the first two terms. Further, the second term in~\eqref{eq:decomposed} is independent of $w^{\perp}$. Finally, the first term will always increase for any non-zero value of $w^{\perp}$ as $||w^{\perp}+w^{||}||^{2} \geq ||w^{||}||$ for any $w^{\perp}\neq0$.

@@

      \paragraph{Bag-of-visual-words.}
         \textcolor{petr}{
            For bag of visual words image representation we provide three kind of results summarized in table \ref{tab:recallBOW}. First, similarly to \cite{Gronat13}, we compare the proposed method (e-SVM) to the standard bag-of-visual-words baseline (BOW). We observe improvement from 28.7\% recall@K=1 (BOW) to 31.8\% recall@K=1 (BOW e-SVM). Secondly, the results are also confronted to the p-value calibration of \cite{Gronat13}. The p-value calibration (BOW p-val) gives marginally better results (33.0\% recall@K=1) but at an increased training cost and a significantly higher required memory footprint at query time, which makes the method hard to scale-up to database sizes beyond 200k images. Finally, we investigate the performance of a raw exemplar-SVM (raw e-SVM) with no calibration. As reported in \cite{Gronat13}, we observe that using raw SVM scores directly with no calibration does not perform on BOW. However, this method gives satisfiable good results on Fisher vectors. Details are be given next.
         }

         \begin{table}[b]
            \begin{centering}
               \include{table25kBOW}
            \caption{ \textbf{Results on Pittsburgh 25k dataset - BOW.}
                  The fraction of correctly recognized queries (recall@K)
                  \textcolor{petr}{
                     in the top $K\in\{1,2,5,10,20\}$ retrieved images.
                     % Different methods have been applied to two types of descriptors, the BOW and Fisher vectors compressed to different dimensions. The learnt representations (BOW \emph{e-SVM} and BOW \emph{p-val}) outperform the raw BOW baseline as well as the learnt representation without calibration (BOW raw SVM).
                  } 
                  The learnt descriptors by the proposed method (BOW e-SVM) marginally improves over the raw bag-of-visual-words descriptors (BOW).
%                   on both the 25k and 55k Pittsburgh image datasets.     
            }
            \label{tab:recallBOW}
            \end{centering}
         \end{table}

      \paragraph{Fisher vectors.}

@@

         Figure \ref{fig:recall} shows the results in a form of curve and it clearly demonstrates the benefits of the learnt descriptors. For comprehensiveness the figure also shows the BOW baseline curve taken from table \ref{tab:recallBOW} shown in black.

@@

         The p-val calibration of~\cite{Gronat13} has been also tested on e-SVM Fisher vectors. We found that the calibration did not perform well for the learnt Fisher vector classifiers (top 1 recall of 25.3\% compared to baseline performance of 33.6\% for dimension 128). When examining the results, we believe this may be attributed to the fact that Fisher vectors are low dimensional compared to the extremely high-dimensional (d=100,000) bag-of-visual-words representation, which affects the distribution of the classifier scores on the negative test data.

@@

         Next, we investigate the performance of a raw exemplar-SVM (raw e-SVM) without any further calibration on Fisher vectors. We observe that while using raw SVM scores directly without any calibration fails for the bag-of-visual-words representation (BOW raw SVM, talbe~\ref{tab:recallBOW}), as reported in \cite{Gronat13}, it marginally improves the recall for larger $K$ for Fisher vectors (68.4\% for FV128 raw SVM compared to 67.7\% for non-learned FV128).

@@

         Finally, we performed experiments on the larger Pittsburgh 55k dataset. Similarly to the former dataset, the results summarized in table \ref{tab:recall2} show consistent improvement of the proposed method (FV e-SVM) over the Fisher vector baseline (FV) across all target dimensions. The benefits of the proposed method in terms of memory efficiency will be discussed next.
